Synthesis of Rules for Network Intrusion Detectors

### 2019-10-24 paper/paper.tex
pass on rules and packets

\Fix{Illustrate how information is extracted from packets}

@@

Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
this attack. This rule was obtained from the \Fix{...what source...}
The relevant options in the rule appear in bold---the option
\CodeIn{flags: S,12}, identifies a SYN packet in a TCP packet and the
option \CodeIn{threshold: type both, track by\_dst, count 5000,
  seconds 5} indicates that a high volume of such packets should be
requested in a short period of time.

@@

\Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
\begin{figure}[h]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

@@

\begin{wrapfigure}[14]{r}{0.56\textwidth}
  \centering
  \vspace{-8ex}
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule (up to 10).}
\end{wrapfigure}
The goal of the first step is to extract options from the positive
traffic. For that, \tname{} parses the protocol message, looking for
fields in the message associated with options (see
Section~\ref{sec:rules-and-packets}). For example, fields
\Fix{...and...}  can be used to determine the option
\Fix{zzz}. Unfortunately, inferring options at the field granularity
level is insufficient to capture \CodeIn{content} options (see
Section~\ref{sec:example-suricata-rules}), which refer to parts of the
payload of the message and cannot be directly extracted directly from
the fields. Handling this case is critically important as most rules
contain \CodeIn{content}
options. Figure~\ref{fig:distribution-contents} shows the histogram of
number of these options per rule (up to 10) for the Suricata public
rule database \Fix{cite}. Rules without \CodeIn{content} options is an
exception and half of the rules contain at least two \CodeIn{content}
options.  To find the values associated with these options, \tname{}
splits the payload of the malicious message in tokens using natural
language delimiters, such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. The rationale for this decision is that we
observed, by inspecting existing rules, that \CodeIn{content} options
typically refer to sequence of characters in the payload of messages
separated by these delimiters. This tokenization step can generate
dozens of options. For the example from
Section~\ref{sec:content-example}, \tname{} produces a total of
\Fix{xx} content options.

>>>>

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.3]{figs/ICMP-packet-structure.png}
\caption{ICMP Packet Structure}
\label{fig:overview}
\end{figure*}

\Luc{Os campos Flags e Time to live no header IP são mapeados para as opções fragbits e ttl, respectivamente. Os campos Type e Code no header ICMP são mapeados para as opções itype e icode. \\
Os campos, Time to Live, Type e Code recebem diretamente o valor que está no pacote. Por exemplo, se o campo Type no pacote possuir valor 8, na regra ele vai aparecer como itype:8.\\
O campo Flags é intepretado como um vetor de bits, onde cada posição representa uma flag. Na regra, o valor do campo Flags é traduzido em caracteres que representam cada flag presente no pacote. Por exemplo: se o campo Flags possuir o valor 1010, na regra aparecerá fragbits:RM, pois os caracteres R e M indicam a presença do quarto e segundo bits.}

@@

Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
this attack. This rule was obtained from the \Luc{Emerging Threats Open Ruleset}.
The relevant options in the rule appear in bold---the option
\CodeIn{flags: S,12}, identifies a SYN packet in a TCP packet and the
option \CodeIn{threshold: type both, track by\_dst, count 5000,
  seconds 5} indicates that a high volume of such packets should be
requested in a short period of time.

@@

\begin{figure}[h]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

@@

\begin{wrapfigure}[14]{r}{0.56\textwidth}
  \centering
  \vspace{-8ex}
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule (up to 10).}
\end{wrapfigure}
The goal of the first step is to extract options from the positive
traffic. For that, \tname{} parses the protocol message, looking for
fields in the message associated with options (see
Section~\ref{sec:rules-and-packets}). For example, fields
\Fix{...and...}  can be used to determine the option
\Fix{zzz}. Unfortunately, inferring options at the field granularity
level is insufficient to capture \CodeIn{content} options (see
Section~\ref{sec:example-suricata-rules}), which refer to parts of the
payload of the message and cannot be directly extracted directly from
the fields. Handling this case is critically important as most rules
contain \CodeIn{content}
options. Figure~\ref{fig:distribution-contents} shows the histogram of
number of these options per rule (up to 10) for the \Luc{Emerging Threats Open Ruleset}. Rules without \CodeIn{content} options is an
exception and half of the rules contain at least two \CodeIn{content}
options.  To find the values associated with these options, \tname{}
splits the payload of the malicious message in tokens using natural
language delimiters, such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. The rationale for this decision is that we
observed, by inspecting existing rules, that \CodeIn{content} options
typically refer to sequence of characters in the payload of messages
separated by these delimiters. This tokenization step can generate
dozens of options. For the example from
Section~\ref{sec:content-example}, \tname{} produces a total of
\Fix{xx} content options.
